---
title: 'PS 5625 PS #6 Development File'
author: "Patrick Edwards"
date: "5/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

PRELIMINARY: Load package `Rcpp`:
```{r}
#install.packages("Rcpp")
library(Rcpp)
```

# PRELIMINARY: Start with assumption that all models are weighted equally.

## Load in test data.
```{r}
library(EBMAforecast)
df <- EBMAforecast::presidentialForecast
y <- df$Actual
```


## R code.
```{r eval = FALSE}
## Find number of models:
K <- dim(df)[2] - 1

## Find the number of observations (in this case, presidential elections):
T <- dim(df)[1]

## Define w_hat with all models weighted equally.
w_hat <- replicate(K,1/K)
```

# STEP (1): Write an Rcpp function that will calculate the answer to Equation (2). The output will be a matrix.

## Corresponding R Code (From in-class assignment)
```{r eval = FALSE}
z_hat <- (w_hat * dnorm(y[t], 
                        df[t,k], 
                        1
                        )
          )/(sum(w_hat * dnorm(y[t], df[t,k], 1)))
```

## Rcpp Code

**NOTE**: this is largely from `Rcpp` code that my group created for in-class assignment #21.

```{r rcppVersion}
library(Rcpp)
library(RcppArmadillo)
# Code for Equation 2:
## NOTE: I base this code on in-class assignment #21. Our group created the cpp code that solved step (1) & step(2)

# Baseline example:

#include <Rcpp.h>
using namespace Rcpp;


//' Z-Hat/Computation: Predicting Z Values.
//'
//' @param x Numeric matrix of model predictions.
//' @param y Vector of actual (observed) outcomes for each observation.
//' @param W_H Vector of weights with each element corresponding to a model.
//' @param sd Standard deviation.
//'
//' @export

// [[Rcpp::export]]


NumericMatrix Z_H(NumericMatrix x, 
                  NumericVector y, 
                  NumericVector W_H, 
                  double sd
){
  
  //Start here, creates matrices
  //Code based on inclass assingment
  int rows = x.nrow();
  NumericMatrix dNormal(x);
  NumericMatrix out(x);
  
  //Calculates dnorm like for the R portion of inclass21:
  for (int l = 0; l < x.nrow(); ++l) {
    for (int k = 0; k < x.ncol(); ++k) {
      double dnorm_v = R::dnorm(
        y[l], 
         x(l,k), 
         sd, 
         FALSE
      );
      dNormal(l,k) =  dnorm_v; // 
    }
  }
  
  //weights rows before the summation. 
  //Parallels the summation in the denominator in equation
  NumericVector val(rows);
  for (int l = 0; l < rows; ++l) {
    
    double rs = 0;
    for (int k = 0; k < dNormal.ncol(); ++k) {
      rs += W_H[k] * dNormal(
        l,
        k
      );
    }
    val[l] = rs;
  }
  
  
  // Calculates Z_H value.
  for (int l = 0; l < x.nrow(); ++l) {
    for (int k = 0; k < x.ncol(); ++k) {
      out(l,k) =  W_H[k] * dNormal(l,k) / val[l]; // 
    }
  }
  
  // Equivalent to return() in R. Returns matrix of z-hats
  return out;
}



```


# Step (2):

```{r}
//' W-Hat/Reweighting Computation.
//'
//'
//' @param Z_H Numeric matrix of z-hat values, derived in Z_H function
//'
//' @export


// [[Rcpp::export]]

// Reweighting function, similar to what was done on our in-class assignment.
NumericVector W_H(NumericMatrix Z_H){
  
  // Derive object for # of Z_hat rows:
  int rows = Z_H.nrow();
  
  // Derive object for # of Z_hat columns:
  int cols = Z_H.ncol();
  NumericVector out(cols);
  
  // Reweighting function. Code based on W-hat calculations done 
  //  for inclass assignment #21.
  for (int k = 0; k < Z_H.ncol(); ++k) {
    
    double nWeight = 0;
    
    // Inner for-loop sums the z-hats column-wise:
    for (int i = 0; i < Z_H.nrow(); ++i) {
      nWeight +=  Z_H(i,k);
    }
    
    //Column-wise sum of Z-hats are divided by the # of Z-hat rows.
    weight_out[k] = nWeight / rows;
  }
  
  // Returns new weights:
  return weight_out;
}


```




# STEP (3): Write an Rcpp function that will complete the entire algorithm.

## Corresponding R Code (From in-class assignment)

```{r eval = FALSE}
# PRELIMINARY 1: Number of models:
K <- dim(df)[2] - 1

# PRELIMINARY 2: Number of observations:
Ti <- dim(df)[1]

# PRELIMINARY 3: Create function that combines steps (2) & (3):
steps23 <- function(df = df, w_hat = wHat_it[i-1,]){
  y <- df[[7]] # Create vector of outcome variables.
  M <- df[,1:K] # create matrix of models.
  K <- dim(df)[2] - 1 # Number of models.
  Ti <- dim(df)[1] #Number of observations
  # Create dnorm object for step 2:
  dnormal <- t(
    sapply(
      X = 1:Ti,
      FUN = function(t){
        dnorm(
          y[t],
          sapply(
            X = 1:K,
            FUN = function(k){df[t,k]}
            ),
          sd = 1
          )
        }
      )
    )
  # Do step 2 function:
  z_hats <- t(
    sapply(
      X = 1:Ti,
      FUN = function(t){
        (w_hat * dnormal[t,])/(sum(w_hat * dnormal[t,]))
        }
      )
    )
  # Do step 3 function:
  w_hat2 <- sapply(
    X = 1:K,
    FUN = function(k){
      sum(z_hats[,k])/Ti
    }
  )
  return(w_hat2)
}

# PRELIMINARY 4: create matrix to collect iterations data. It needs to have 21
# rows for the initial w_hat and the 20 iterations of w_hat.
itnum <- 20 # Number of iterations
wHat_it <- matrix(
  data = NA,
  nrow = itnum + 1,
  ncol = K
  )
colnames(wHat_it) <- c("Campbell",
                       "Lewis-Beck",
                       "EWT2C2",
                       "Fair",
                       "Hibbs",
                       "Abramowitz"
                       )
rownames(wHat_it) <- 0:itnum

# STEP 1: Define w_hat with all models weighted equally.
wHat_it[1,] <- replicate(K,1/K)

# Steps 2-4 (ITERATION):
i <- 2 
while(i <= itnum + 1){
  wHat_it[i,] <- steps23(df = df, w_hat = wHat_it[i-1,])
  print(i)
  i = i + 1
}

# FINAL ITERATION:
wHat_it[tail(itnum,1),]
```

## Rcpp Code

```{r rcppVersion}
# ALso loosly based on in-class assignment.



//' EBMA Calculation Using C++
//'
//' @param x Numeric matrix of  predictions from each model.
//' @param y Numeric vector of actual (observed) outcomes.
//' @param W_Hats Numeric vector of the weight assigned to each model.
//' @param sd Standard deviation.
//' @param tol Tolerance threshold after which iteration ends.
//'
//' @export
// [[Rcpp::export]]

// Please note that this function is loosely based off the thre

NumericVector CPP_EBMA(
    NumericMatrix x, 
    NumericVector y, 
    NumericVector W_Hats, 
    double sd, 
    double tol
  ){
  
  // If tolerance parameter is met, then the iteration does not run.
  bool threshold = FALSE;
  int iterations = 0;
  
  // Derive Z-hats as well as updated weights W-hats:
  while(threshold == FALSE){
    
    NumericMatrix C_Z_H = Z_H(
      x, 
      y, 
      W_Hats, 
      sd
    );
    
    NumericVector nWeights = W_H(C_Z_H);
    
    // This determines if the tolerance parameter has been met. 
    // TRUE if all elements in weights object fall below tol threshold.
    // If TRUE, then ends the iterated process and prints the total number of iterations
    
    LogicalVector diff = (abs((W_Hats - nWeights)) < tolerance);
    
    if(is_true(all(diff))){
      threshold = TRUE;
    }
    
    else{
      W_Hats = nWeights;
    }
    
    iterations = iterations + 1;
  
  }
  
  return W_Hats;
}


```



# STEP (4): Include one unit test per function.

**[STEP NOT COMPLETE]**


# STEP (5): Assemble the code as an R package.

```{r}
library(Rcpp)
#Rcpp.package.skeleton("rcppEBMA")
#usethis::use_rcpp("rcppEBMA")


#getwd()
#setwd("~/")
#library(usethis)
#usethis::use_rcpp("C:/Users/edwar/Documents/RcppEBMA")


Rccp::compile_attributes()


# Add three steps


# Create outside of github repository, then put it back in later.
#setwd("Documents")
#library(usethis)
#usethis::use_rcpp("RcppEBMA")
```


# STEP (6): Write a development .R file that demonstrates the package’s use.

**[STEP NOT COMPLETE]**




