---
title: 'PS 5625 PS #6 Development File'
author: "Patrick Edwards"
date: "5/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

PRELIMINARY: Load package `Rcpp`:
```{r}
#install.packages("Rcpp")
library(Rcpp)
```

# PRELIMINARY: Start with assumption that all models are weighted equally.

## R code.
```{r eval = FALSE}
## Find number of models:
K <- dim(df)[2] - 1

## Find the number of observations (in this case, presidential elections):
T <- dim(df)[1]

## Define w_hat with all models weighted equally.
w_hat <- replicate(K,1/K)
```

## Rcpp Code (is this necessary?)

**[NOT DONE]**


# STEP (1): Write an Rcpp function that will calculate the answer to Equation (2). The output will be a matrix.

## Corresponding R Code (From in-class assignment)
```{r eval = FALSE}
z_hat <- (w_hat * dnorm(y[t], df[t,k], 1))/(sum(w_hat * dnorm(y[t], df[t,k], 1)))
```

## Rcpp Code

**NOTE**: this is largely from `Rcpp` code that my group created for in-class assignment #21.

```{r}
library(Rcpp)

# Deriving Eq. 2:

cppFunction('
  NumericMatrix cpp_Eq2(
    NumericMatrix x,
    NumericVector y,
    NumericVector weights,
    double sd
    ){
      int rows = x.nrow();
      
    }
  
')
```


# STEP (2): Write an Rcpp function that will calculate the answer to Equation (3). The output will be a vector.

## Corresponding R Code (From in-class assignment)

```{r eval = FALSE}
## NOTE 1: input `z_hat` and output `w_hat` are vectors of `K` elements.
## NOTE 2: `T = dim(df)[1]`, or the number of all presidential elections in sample.
w_hat <- sum(z_hat)/T
```

## Rcpp Code

**[NOT DONE]**


# STEP (3): Write an Rcpp function that will complete the entire algorithm.

## Corresponding R Code (From in-class assignment)

```{r eval = FALSE}
# PRELIMINARY 1: Number of models:
K <- dim(df)[2] - 1

# PRELIMINARY 2: Number of observations:
Ti <- dim(df)[1]

# PRELIMINARY 3: Create function that combines steps (2) & (3):
steps23 <- function(df = df, w_hat = wHat_it[i-1,]){
  y <- df[[7]] # Create vector of outcome variables.
  M <- df[,1:K] # create matrix of models.
  K <- dim(df)[2] - 1 # Number of models.
  Ti <- dim(df)[1] #Number of observations
  # Create dnorm object for step 2:
  dnormal <- t(
    sapply(
      X = 1:Ti,
      FUN = function(t){
        dnorm(
          y[t],
          sapply(
            X = 1:K,
            FUN = function(k){df[t,k]}
            ),
          sd = 1
          )
        }
      )
    )
  # Do step 2 function:
  z_hats <- t(
    sapply(
      X = 1:Ti,
      FUN = function(t){
        (w_hat * dnormal[t,])/(sum(w_hat * dnormal[t,]))
        }
      )
    )
  # Do step 3 function:
  w_hat2 <- sapply(
    X = 1:K,
    FUN = function(k){
      sum(z_hats[,k])/Ti
    }
  )
  return(w_hat2)
}

# PRELIMINARY 4: create matrix to collect iterations data. It needs to have 21
# rows for the initial w_hat and the 20 iterations of w_hat.
itnum <- 20 # Number of iterations
wHat_it <- matrix(
  data = NA,
  nrow = itnum + 1,
  ncol = K
  )
colnames(wHat_it) <- c("Campbell",
                       "Lewis-Beck",
                       "EWT2C2",
                       "Fair",
                       "Hibbs",
                       "Abramowitz"
                       )
rownames(wHat_it) <- 0:itnum

# STEP 1: Define w_hat with all models weighted equally.
wHat_it[1,] <- replicate(K,1/K)

# Steps 2-4 (ITERATION):
i <- 2 
while(i <= itnum + 1){
  wHat_it[i,] <- steps23(df = df, w_hat = wHat_it[i-1,])
  print(i)
  i = i + 1
}

# FINAL ITERATION:
wHat_it[tail(itnum,1),]
```

## Rcpp Code

**[NOT DONE]**


# STEP (4): Include one unit test per function.

**[STEP NOT COMPLETE]**


# STEP (5): Assemble the code as an R package.

**[STEP NOT COMPLETE]**


# STEP (6): Write a development .R file that demonstrates the package’s use.

**[STEP NOT COMPLETE]**




